\begin{abstract}

We set out to design a system for performing trusted computation on mainstream
desktop and server-class computers. Our system has three mutually distrusting
parties: \textit{a data provider} has some private information, and allows
\textit{a software provider} to perform some computation on the private data,
using hardware resources provided by \textit{an infrastructure owner}. The
computation result is communicated to either the data provider or the software
provider, in encrypted form.

We aim to build a system that can prevent malicious software from colluding
with either malicious infrastructure, or with another piece of malicious
software running on the same infrastructure, for the purpose of leaking private
data that is not included in the computation result. Our trusted computing base
(TCB) includes the computer's main processor (CPU), and a
\textit{software loader} that verifies and launches the untrusted software.
We rely on a processor's ability to create a secure execution environment that
is isolated from the system software. Our TCB does not include the operating
system or any high-level software that runs on the infrastructure.

The main application that we are targeting is enabling heavily regulated
fields, such as the medical and financial industries, to use public clouds and
software as a service (SaaS). While legal consequences make it unlikely that a
public infrastructure or SaaS provider will deliberately act in a malicious
manner, our threat model covers against skilled attackers who can exploit bugs
in both the infrastructure and the software that performs computation over the
private data.

We have analyzed Intel's most recent proposal for achieving trusted computing,
Software Guard Extensions (SGX), and find that it leaks private data by
exposing the software's memory access patterns to other programs sharing the
infrastructure and to the infrastructure's owners. We aim to find a minimal set
of changes to the SGX proposal that would make it not reveal memory access
patterns, when used in conjunction with our software loader.

\end{abstract}
